---
title: "Investigating COVID-19 Virus Trends"
author: "Stephan Hertweck"
date: "3 12 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##

The Covid-19 pandemic has confronted the international community with increased demands for disease surveillance of the general population. Widespread testing has become one of the most important tools to monitor disease spread, but variability in testing capacity between countries makes comparison less straightforward.

In this document I aim to compare the country-specific incidence against that nation's testing frequency.

```{r, message=FALSE}
# load data
library(readr)
library(dplyr)

covid_df <- read_csv("covid19_data.csv")

dim(covid_df)

vector_cols <- colnames(covid_df)

head(data)

glimpse(covid_df)

```

# Data preparation
Filter the rows related to "All States" from the Province_State column and remove the Province_State column from covid_df dataframe.

```{r, message=FALSE}
covid_df_all_states <- covid_df %>% filter(Province_State == "All States") %>% select(-Province_State)


# Keep only daily data (rather than cumulative)

covid_df_all_states_daily <- select(covid_df_all_states, Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)

#Sum cases for each country and sort in descending fashion

covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% group_by(Country_Region) %>% summarize(tested = sum(daily_tested), positive = sum(daily_positive), active = sum(active), hospitalized = sum(hospitalizedCurr)) %>% arrange(-tested)

# Top 10

covid_top_10 <-  head(covid_df_all_states_daily_sum, 10)


# Identifying the highest against positive cases 

countries <- covid_top_10$Country_Region
tested_cases <-  covid_top_10$tested
positive_cases <- covid_top_10$positive
active_cases <- covid_top_10$active
hospitalized_cases <-  covid_top_10$hospitalized

names(tested_cases) <- countries
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries

# top 3 countries

ratio <- positive_cases/tested_cases
ratio <-  sort(ratio, decreasing = TRUE)

positive_tested_top_3 <- head(ratio, 3)

# summarize results

united_kingdom <-  c(0.11, 1473672, 166909, 0, 0)
united_states <- c(0.10, 17282363, 1877179, 0, 0)
turkey <- c(0.08, 2031192, 163941, 2980960, 0)

covid_mat <-  rbind(united_kingdom, united_states, turkey)

colnames(covid_mat) <- c("Ratio", "tested", "positive", "active", "hospitalized")

print(covid_mat)


# oraganize summary
dfs <-  list(c(covid_df, covid_df_all_states, covid_df_all_states_daily, covid_top_10))
matrx <-  list(covid_mat)
vectors <-  list(c(vector_cols, countries))

data_structure_list <-  list(c(dataframes = dfs, matrices =  matrx, vectors = vectors))

question <- "Which countries have had the highest number of positive cases against the number of tests?"
answer <-  c("Positive tested cases" = positive_tested_top_3)

covid_analysis_list <- list(question, answer, data_structure_list)

print(covid_analysis_list[2])

````

The results of this quick analysis show that the UK, the US and turkey have had the highest ratio of positive Covid-19 patients to testing frequency.
